Homework 2 - Ju Young Park
========================================================

### Reproduce the results presented in Figure 2 of the following paper: 
# Qian, F., Bolen, C. R., Jing, C., Wang, X., Zheng, W., Zhao, H., et al. (2013). Impaired toll-like receptor 3-mediated immune responses from macrophages of patients chronically infected with hepatitis C virus. Clinical and Vaccine Immunology : CVI, 20(2), 146–155. doi:10.1128/CVI.00530-12

# Getting the data from GEO
```{r, eval=TRUE}
# You need to evaluate this otherwise the packages would not be loaded
# source("http://bioconductor.org/biocLite.R")

library(GEOquery)
library(Biobase)
library(data.table)
library(limma)
```

```{r}
# You need to use a relative path (make sure the directory exists on your machine)
gds <- getGEO("GSE40812", destdir="./Data/")[[1]]
```

# sanitize data
```{r}
pd <- pData(gds)

mm_pd <- pd[pd$source_name_ch1=="Monocyte-derived Macrophage",]
mm_eset<-gds[,rownames(mm_pd)]
#characteristic_ch1 indicates VL++ or VL--
#characteristic_ch1.2 indicates Mock or Poly IC
mm_pd$HCV <- gsub(".*: ", "", mm_pd$characteristics_ch1)
mm_pd$HCV <- ifelse(mm_pd$HCV=="Neg", "-", "+")
mm_pd$treatment <- gsub(".*: ", "", mm_pd$characteristics_ch1.2)
pData(mm_eset) <- mm_pd
```

# Normalizing the data
```{r, eval=FALSE}
library(lumi)

mm_eset <- lumiN(mm_eset)
#Note: Expression matrix has been quantile normalized using the lumi package for practice. Normalization step in this data will not change the data. 
```

# Using LIMMA to test for differential expression
```{r, message=FALSE}
# load the limma package
library(limma)
```

```{r}
# test for differential expression using limma
treat_matrix <- model.matrix(~treatment, mm_eset)
colnames(treat_matrix)
colnames(treat_matrix) <- c("Mock", "Poly")
fit1 <- lmFit(mm_eset, treat_matrix)
ebay1 <- eBayes(fit1)

# Top table of results, adjusted for false discovery rate (5%)
TopTable1 <- topTable(ebay1, coef="Poly", p.value=0.05, lfc=log2(1.5),number=Inf)

selected <- TopTable1$ID
length(selected)
```

```{r}
eset_small <- mm_eset[selected,]
HCV_matrix <- model.matrix(~HCV, eset_small)
colnames(HCV_matrix)
colnames(HCV_matrix) <- c("Neg", "Pos")
fit7 <- lmFit(eset_small, HCV_matrix)
fit7_ebay <- eBayes(fit7)
TopTable7 <- topTable(fit7_ebay, coef="Pos", adjust="none", number=Inf, p.value=0.1)
selected <- TopTable7$ID
length(selected)
```


# Results using a heatmap (Use the pheatmap package)
```{r}
library(pheatmap)
mat <- exprs(eset_small[as.character(selected),])
pheatmap(mat, scale="row")
```
