Homework 2 - Ju Young Park
========================================================

### Reproduce the results presented in Figure 2 of the following paper: 
# Qian, F., Bolen, C. R., Jing, C., Wang, X., Zheng, W., Zhao, H., et al. (2013). Impaired toll-like receptor 3-mediated immune responses from macrophages of patients chronically infected with hepatitis C virus. Clinical and Vaccine Immunology : CVI, 20(2), 146–155. doi:10.1128/CVI.00530-12

# Getting the data from GEO
```{r, eval=FALSE}
source("http://bioconductor.org/biocLite.R")

library(GEOquery)
library(Biobase)
library(data.table)
library(limma)
```

```{r}
gds <- getGEO("GSE40812", destdir="/Users/Julie/Documents/R/BIOST578/DB")

#Extract data file mannually
gds <- getGEO(filename = "/Users/Julie/Documents/R/BIOST578/DB/GSE40812_series_matrix.txt")
```

# sanitize data
```{r}
pd <- pData(gds)

sanitize_pdata <- function(pd){
  keepCols <- c(
    "characteristics_ch1", "characteristics_ch1.2")
  pd <- pd[, keepCols]
  colnames(pd) <- c("ptkinds", "treatment")
  pd$ptkinds <- gsub(".*: ", "", pd$ptkinds)
  pd$treatment<-gsub("treatment: ", "", pd$treatment)
  pd
}

mm_pd <- pd[pd$source_name_ch1=="Monocyte-derived Macrophage",]
mm_eset<-gds[,rownames(mm_pd)]
#characteristic_ch1 indicates VL++ or VL--
#characteristic_ch1.2 indicates Mock or Poly IC
mm_pd$HCV <- gsub(".*: ", "", mm_pd$characteristics_ch1)
mm_pd$HCV <- ifelse(mm_pd$HCV=="Neg", "-", "+")
mm_pd$treatment <- gsub(".*: ", "", mm_pd$characteristics_ch1.2)

pData(gds) <- sanitize_pdata(pData(gds))
```

# NOrmalizing the data
```{r}
library(lumi)

gds <- lumiN(gds)
#Note: Expression matrix has been quantile normalized using the lumi package for practice. Normalization step in this data will not change the data. 
```

# Using LIMMA to test for differential expression
```{r, message=FALSE}
# load the limma package
library(limma)
```

```{r}
# test for differential expression using limma
treat_matrix <- model.matrix(~treatment+0, mm_pd)
colnames(treat_matrix)
colnames(treat_matrix) <- c("Mock", "Poly")
fit1 <- lmFit(mm_eset, treat_matrix)
ebay1 <- eBayes(fit1)

# Top table of results, adjusted for false discovery rate (5%)
TopTable1 <- topTable(ebay1, coef="Poly", number=Inf, p.value=0.05, adjust="BH")
sum(TopTable1$adj.P.Val < 0.1)

selected <- p.adjust(ebay1$p.value[, 3]) <0.1
gds_selected <-gds[[1]][selected,]
```

# Results using a heatmap (Use the pheatmap package)
```{r}
library(pheatmap)

pData(plot)
heatmap(exprs(gds_Selected))
```
